### 2022-06-27 thesis.tex
feat: add hsrm cover page

\documentclass[11pt,
  paper=a4, 
  bibliography=totocnumbered,
	captions=tableheading,
	BCOR=10mm
]{scrreprt}

\usepackage[german]{babel} % deutsch und deutsche Rechtschreibung
\usepackage[T1]{fontenc} % Umlaute und deutsches Trennen
\usepackage{textcomp} % Euro
\usepackage[hyphens]{url}
\usepackage{emptypage} % Wirklich leer bei leeren Seiten

\usepackage{newcent}
\usepackage{microtype}
\usepackage[paper=a4paper,width=14cm,left=35mm,height=22cm]{geometry}
\setlength{\parskip}{0.5em}
\setlength{\parindent}{0em} % im Deutschen Einrückung nicht üblich, leider

% Seitenmarkierungen 
\newcommand{\phv}{\fontfamily{phv}\fontseries{m}\fontsize{9}{11}\selectfont}
\usepackage{fancyhdr} % Schickere Header und Footer
\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
%\fancyhead[L]{\phv \leftmark}
\fancyhead[RE,LO]{\phv \nouppercase{\leftmark}}
\fancyhead[LE,RO]{\phv \thepage}
% Unten besser auf alles Verzichten
%\fancyfoot[L]{\textsf{\small \kurztitel}}
\fancyfoot[C]{\ } % keine Seitenzahl unten
%\fancyfoot[R]{\textsf{\small Medieninformatik}}

\usepackage{hsrmlogo}

\usepackage[utf8]{inputenc}

\usepackage{setspace}
\usepackage{csquotes} % Context sensitive quotation.
\usepackage{amsmath} % Standard math.
\usepackage{amsthm} % Math theorems.
\usepackage{amssymb} % More math symbols.
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]

\usepackage[section]{placeins} % Keep floats in the section they were defined in.
\usepackage{tabularx}
\usepackage{booktabs} % Scientific table styling.
\usepackage{floatrow} % Option for keeping floats in the place they were defined in the code.
\floatsetup[table]{style=plaintop}
\usepackage{hyperref} % Hyperlinks.
\usepackage[all]{nowidow} % Prevent widows and orphans.
\usepackage{xstring} % logic string operations
\usepackage{bbm} % \mathbb on numerals.
\usepackage{csquotes}
\usepackage{mathtools}
\usepackage[ruled,vlined]{algorithm2e} % Pseudocode
\usepackage{scrhack} % Make warning go away.
\usepackage{graphicx}
\usepackage{subcaption} % Subfigures with subcaptions.
\usepackage{authoraftertitle} % Make author, etc., available after \maketitle
\usepackage{listofitems}
\usepackage{blindtext} % Placeholder text.
\usepackage[automake, nopostdot, nonumberlist]{glossaries} % glossary for definitions and acronyms, without dot after entry and page reference 
\makeglossaries % Generate the glossary

% \PassOptionsToPackage{obeyspaces}{url}%
\usepackage[backend=bibtex,% 
style=nature,% 
doi=true,isbn=false,url=false, eprint=false]{biblatex}
% \renewbibmacro*{url}{\printfield{urlraw}}

\addbibresource{references.bib}

\DeclareStyleSourcemap{
  \maps[datatype=bibtex, overwrite=true]{
    \map{
      \step[fieldsource=url, final]
      \step[typesource=misc, typetarget=online]
    }
    \map{
      \step[typesource=misc, typetarget=patent, final]
      \step[fieldsource=institution, final]
      \step[fieldset=holder, origfieldval]
    }
  }
}

\linespread{1.15} % set line spacing

\usepackage{listings} % rendering program code
\lstset{% general command to set parameter(s)
	basicstyle=\ttfamily\color{grey},          % print whole listing small
	keywordstyle=\color{black}\bfseries\underbar,
	% underlined bold black keywords
	identifierstyle=,           % nothing happens
	commentstyle=\color{white}, % white comments
	stringstyle=\ttfamily,      % typewriter type for strings
	showstringspaces=false}     % no special string spaces
\lstset{language=python} % und nur schöne Programmiersprachen ;-)

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
      <5> <6> <7> <8> <9> <10>
      <10.95> <12> <14.4> <17.28> <20.74> <24.88>
      mathx10
      }{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareFontSubstitution{U}{mathx}{m}{n}
\DeclareMathSymbol{\bigtimes}{1}{mathx}{"91}

%%% Custom definitions %%%
% Shorthands
\newcommand{\welchethesis}{Master}
\newcommand{\thesisofwas}{of Science}
\newcommand{\titel}{Data Mining komplexer Datenstrukturen aus PDF-Dokumenten}
\newcommand{\kurztitel}{Data Mining PDF-Dokumenten}
\newcommand{\autor}{Deniz Aydar}
\newcommand{\datum}{25. Juli 2022} % Abgabedatum
\newcommand{\ort}{Wiesbaden}
\newcommand{\referent}{Prof.\ Dr.\ Dirk Krechel}
\newcommand{\korreferent}{Prof.\ Dr.\ Philipp Schaible}

\newcommand{\ie}{i.\,e.~}
\newcommand{\eg}{e.\,g.~}
\newcommand{\ind}{\mathbbm{1}}
% Functions
\newcommand{\tpow}[1]{\cdot 10^{#1}}
\newcommand{\figref}[1]{(Figure \ref{#1})}
\newcommand{\figureref}[1]{Figure \ref{#1}}
\newcommand{\tabref}[1]{(Table \ref{#1})}
\newcommand{\tableref}[1]{Table \ref{#1}}
\newcommand{\secref}[1]{%
	\IfBeginWith{#1}{chap:}{%
		(cf. Chapter \ref{#1})}%
		{(cf. Section \ref{#1})}%
		}
\newcommand{\sectionref}[1]{%
	\IfBeginWith{#1}{chap:}{%
		Chapter \ref{#1}}%
		{\IfBeginWith{#1}{s}{%
			Section \ref{#1}}%
			{[\PackageError{sectionref}{Undefined option to sectionref: #1}{}]}}}
\newcommand{\chapref}[1]{(see chapter \ref{#1})}
\newcommand{\unit}[1]{\,\mathrm{#1}}
\newcommand{\unitfrac}[2]{\,\mathrm{\frac{#1}{#2}}}
\newcommand{\codeil}[1]{\lstinline{#1}}{} % wrapper for preventing syntax highlight error
\newcommand{\techil}[1]{\texttt{#1}}
\newcommand{\Set}[2]{%
  \{\, #1 \mid #2 \, \}%
}
% Line for signature.
\newcommand{\namesigdate}[1][5cm]{%
	\vspace{5cm}
	{\setlength{\parindent}{0cm}
	\begin{minipage}{0.3\textwidth}
		\hrule 
		\vspace{0.5cm}
		{\small city, date}
	\end{minipage}
	 \hfill
	\begin{minipage}{0.3\textwidth}
		\hrule
		\vspace{0.5cm}
	    {\small signature}
	\end{minipage}
	}
}
% Automatically use the first sentence in a caption as the short caption.
\newcommand\slcaption[1]{\setsepchar{.}\readlist*\pdots{#1}\caption[{\pdots[1].}]{#1}}

% Variables. 
% Adapt if necessary, use to refer to figures and graphics.
\def \figwidth {0.9\linewidth}
\graphicspath{ {./graphics/figures/}{./graphics/figures/} } % Path to figures and images.

% Customizations of existing commands.
\renewcommand{\vec}[1]{\mathbf{#1}}
% Capitalized \autoref names.
\renewcommand*{\chapterautorefname}{Chapter}
\renewcommand*{\sectionautorefname}{Section}

% TODO Fill with your data.
\title{My full title}
\author{Firstname Lastname}

\begin{document}

\include{vorspann} % Titelseite, Erklärungen, etc.

\begin{abstract}

	In der Welt der berufsbedingten Mobilität können digitale Lösungsansätze die Arbeit erleichtern.
	So auch in der Personenschifffahrt, bei der neben den üblichen Aufgaben der Arbeit Hindernisse unterschiedlicher Art auftreten können.
	Treten dort weitere Probleme wie das Fehlen einer Internetverbindung auf, kann die digitale Lösung nur bedingt weiterhelfen.
	Damit der Kapitän dies bei der Fahrt und bei der Anlegung vermeidet, bedarf es eine Webanwendung, die auch einen Netzwerkverlust handhaben kann und dabei funktionsfähig bleibt.
	So soll die Anwendung dem Kapitän erlauben, während den Fahrten die Anlegezeiten zu betrachten und die Betankung von Trinkwasser an den Schiffsanleger interaktiv verfolgen zu können.
	Schwerpunkt in dieser Arbeit ist daher die Verarbeitung der Ressourcen an den Anlegestellen am Beispiel der Landebrücken und die Aufarbeitung mit der Problematik des Offline-Zustandes.

\end{abstract}

\tableofcontents
\listoffigures
\listoftables
\listofalgorithms

\chapter{Introduction}
\pagenumbering{arabic}
\blindtext

\begin{figure}[H]
	\centering
	\includegraphics[width=\figwidth]{scientific_paper_graph_quality}
	\slcaption{
		Developmemt of scientific paper graph quality. A dip in the
		quality of scientific graphs is observed from the early 1990s to the early 2010s.
		During this time Microsoft Paint and PowerPoint were often used to create graphs in scientific papers.\label{fig:scientific_graph_quality}}
\end{figure}

\begin{table}[H]
	\begin{tabular}{@{}ll@{}}
		\toprule
		year & quality \\ \midrule
		1985 & good    \\
		2000 & bad     \\ \midrule
		2015 & better  \\ \bottomrule
	\end{tabular}
	\caption{
		Empirical measurements of scientific graph quality. Data points were collected using
		a systematic literature review.\label{tab:scientific_graph_quality}}
\end{table}

This references a \figref{fig:scientific_graph_quality} while this references a table \tabref{tab:scientific_graph_quality}.

A citation looks like this \cite{hadash2018estimate}. To embed a citation in the text flow use textcite,
\eg \textcite{hadash2018estimate} said you should use a lot of citations.

\chapter{Methods}

\chapter{Results and Discussion}

\chapter{Conclusion}

\chapter*{Acknowledgements}
%TODO A place to say thank you to everybody who helped you.

% Acronym definitions
%TODO Add acronym definitions produced by acronyms2glossary.py

\glsaddall
\printglossaries

\printbibliography

\end{document}